Forecasting interrupted time series

### 2024-02-18 fits.tex
Recompiled

\documentclass[11pt,a4paper,]{article}
\usepackage{lmodern}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{
            pdftitle={Forecasting interrupted time series},
            pdfkeywords={blah, blah},
            colorlinks=true,
            linkcolor=blue,
            citecolor=Blue,
            urlcolor=Blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{geometry}
\geometry{left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm}
\usepackage[style=authoryear-comp,]{biblatex}
\addbibresource{references.bib}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\title{Forecasting interrupted time series}

%% MONASH STUFF

%% CAPTIONS
\RequirePackage{caption}
\DeclareCaptionStyle{italic}[justification=centering]
 {labelfont={bf},textfont={it},labelsep=colon}
\captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
\captionsetup[table]{style=italic,format=hang,singlelinecheck=true}

%% FONT
\RequirePackage{bera}
\RequirePackage[charter,expert]{mathdesign}
\RequirePackage[scale=0.9]{sourcecodepro}
\RequirePackage[lf,t]{FiraSans}

%% HEADERS AND FOOTERS
\RequirePackage{fancyhdr}
\pagestyle{fancy}
\rfoot{\Large\sffamily\raisebox{-0.1cm}{\textbf{\thepage}}}
\makeatletter
\lhead{\textsf{\expandafter{\@title}}}
\makeatother
\rhead{}
\cfoot{}
\setlength{\headheight}{15pt}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}
\fancypagestyle{plain}{%
\fancyhf{} % clear all header and footer fields
\fancyfoot[C]{\sffamily\thepage} % except the center
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}}

%% MATHS
\RequirePackage{bm,amsmath}
\allowdisplaybreaks

%% GRAPHICS
\RequirePackage{graphicx}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.8}

%\RequirePackage[section]{placeins}

%% SECTION TITLES
\RequirePackage[compact,sf,bf]{titlesec}
\titleformat{\section}[block]
  {\fontsize{15}{17}\bfseries\sffamily}
  {\thesection}
  {0.4em}{}
\titleformat{\subsection}[block]
  {\fontsize{12}{14}\bfseries\sffamily}
  {\thesubsection}
  {0.4em}{}
\titlespacing{\section}{0pt}{*5}{*1}
\titlespacing{\subsection}{0pt}{*2}{*0.2}

%% TITLE PAGE
\def\Date{\number\day}
\def\Month{\ifcase\month\or
 January\or February\or March\or April\or May\or June\or
 July\or August\or September\or October\or November\or December\fi}
\def\Year{\number\year}

\makeatletter
\def\wp#1{\gdef\@wp{#1}}\def\@wp{??/??}
\def\jel#1{\gdef\@jel{#1}}\def\@jel{??}
\def\showjel{{\large\textsf{\textbf{JEL classification:}}~\@jel}}
\def\nojel{\def\showjel{}}
\def\addresses#1{\gdef\@addresses{#1}}\def\@addresses{??}
\def\cover{{\sffamily\setcounter{page}{0}
        \thispagestyle{empty}
        \placefig{2}{1.5}{width=5cm}{_extensions/numbats/wp/monash2}
        \placefig{16.9}{1.5}{width=2.1cm}{_extensions/numbats/wp/MBSportrait}
        \begin{textblock}{4}(16.9,4)ISSN 1440-771X\end{textblock}
        \begin{textblock}{7}(12.7,27.9)\hfill
        \includegraphics[height=0.7cm]{_extensions/numbats/wp/AACSB}~~~
        \includegraphics[height=0.7cm]{_extensions/numbats/wp/EQUIS}~~~
        \includegraphics[height=0.7cm]{_extensions/numbats/wp/AMBA}
        \end{textblock}
        \vspace*{2cm}
        \begin{center}\Large
        Department of Econometrics and Business Statistics\\[.5cm]
        \footnotesize http://monash.edu/business/ebs/research/publications
        \end{center}\vspace{2cm}
        \begin{center}
        \fbox{\parbox{14cm}{\begin{onehalfspace}\centering\Huge\vspace*{0.3cm}
                \textsf{\textbf{\expandafter{\@title}}}\vspace{1cm}\par
                \LARGE\@author\end{onehalfspace}
        }}
        \end{center}
        \vfill
                \begin{center}\Large
                \Month~\Year\\[1cm]
                Working Paper \@wp
        \end{center}\vspace*{2cm}}}
\def\pageone{{\sffamily\setstretch{1}%
        \thispagestyle{empty}%
        \vbox to \textheight{%
        \raggedright\baselineskip=1.2cm
     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}
        \vspace{2cm}\par
        \hspace{1cm}\parbox{14cm}{\sffamily\large\@addresses}\vspace{1cm}\vfill
        \hspace{1cm}{\large\Date~\Month~\Year}\\[1cm]
        \hspace{1cm}\showjel\vss}}}
\def\blindtitle{{\sffamily
     \thispagestyle{plain}\raggedright\baselineskip=1.2cm
     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}\vspace{1cm}\par
        }}
\def\titlepage{{\cover\newpage\pageone\newpage\blindtitle}}

@@

%% SPACING
\RequirePackage{setspace}
\spacing{1.5}

%% LINE AND PAGE BREAKING
\sloppy
\clubpenalty = 10000
\widowpenalty = 10000
\brokenpenalty = 10000
\RequirePackage{microtype}

%% PARAGRAPH BREAKS
\setlength{\parskip}{1.4ex}
\setlength{\parindent}{0em}

%% HYPERLINKS
\RequirePackage{xcolor} % Needed for links
\definecolor{darkblue}{rgb}{0,0,.6}
\RequirePackage{url}

\makeatletter
\@ifpackageloaded{hyperref}{}{\RequirePackage{hyperref}}
\makeatother
\hypersetup{
     citecolor=0 0 0,
     breaklinks=true,
     bookmarksopen=true,
     bookmarksnumbered=true,
     linkcolor=darkblue,
     urlcolor=blue,
     citecolor=darkblue,
     colorlinks=true}

%% KEYWORDS
\newenvironment{keywords}{\par\vspace{0.5cm}\noindent{\sffamily\textbf{Keywords:}}}{\vspace{0.25cm}\par\hrule\vspace{0.5cm}\par}

%% ABSTRACT
\renewenvironment{abstract}{\begin{minipage}{\textwidth}\parskip=1.4ex\noindent
\hrule\vspace{0.1cm}\par{\sffamily\textbf{\abstractname}}\newline}
  {\end{minipage}}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage[showonlyrefs]{mathtools}
\usepackage[no-weekday]{eukdate}

%% BIBLIOGRAPHY

\makeatletter
\@ifpackageloaded{biblatex}{}{\usepackage[style=authoryear-comp, backend=biber, natbib=true]{biblatex}}
\makeatother
\ExecuteBibliographyOptions{bibencoding=utf8,minnames=1,maxnames=3, maxbibnames=99,dashed=false,terseinits=true,giveninits=true,uniquename=false,uniquelist=false,doi=false, isbn=false,url=true,sortcites=false, date=year}

\DeclareFieldFormat{url}{\texttt{\url{#1}}}
\DeclareFieldFormat[article]{pages}{#1}
\DeclareFieldFormat[inproceedings]{pages}{\lowercase{pp.}#1}
\DeclareFieldFormat[incollection]{pages}{\lowercase{pp.}#1}
\DeclareFieldFormat[article]{volume}{\mkbibbold{#1}}
\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
\DeclareFieldFormat[article]{title}{\MakeCapital{#1}}
\DeclareFieldFormat[inproceedings]{title}{#1}
\DeclareFieldFormat{shorthandwidth}{#1}
% No dot before number of articles
\usepackage{xpatch}
\xpatchbibmacro{volume+number+eid}{\setunit*{\adddot}}{}{}{}
% Remove In: for an article.
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{%
  \printtext{\bibstring{in}\intitlepunct}}}

\makeatletter
\DeclareDelimFormat[cbx@textcite]{nameyeardelim}{\addspace}
\makeatother
\renewcommand*{\finalnamedelim}{%
  %\ifnumgreater{\value{liststop}}{2}{\finalandcomma}{}% there really should be no funny Oxford comma business here
  \addspace\&\space}

\nocover

\author{Rob J~Hyndman, Bahman~Rostami-Tabar}
\addresses{\textbf{Rob J Hyndman}\newline
Department of Econometrics \& Business Statistics\newline
Monash University\newline
Clayton VIC 3800\newline
Australia\newline
{Email: Rob.Hyndman@monash.edu}\newline Corresponding author\newline\\[0.5cm]
\textbf{Bahman Rostami-Tabar}\newline
Cardiff Business School\newline
Wales CF10 3EU\newline
United Kingdom\newline
{Email: rostami-tabarb@cardiff.ac.uk}\\[0.5cm]
}

\date{\sf\Date~\Month~\Year}
\makeatletter
 \lfoot{\sf Hyndman, Rostami-Tabar: \@date}
\makeatother

\usepackage[section]{placeins}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

% Adjust headwidth in case user has changed geometry in header-includes
\renewcommand{\headwidth}{\textwidth}

\begin{document}
\maketitle
\begin{abstract}
A brief summary of our ideas
\end{abstract}
\begin{keywords}
blah, blah
\end{keywords}

\section{Introduction}\label{introduction}

Time series are sometimes interrupted by unusual events; for example, a
natural disaster may occur, or a there may be a temporary policy change.
Many forecasters have faced this issue recently with the COVID-19
pandemic, where historical patterns have been severely disrupted due to
lockdowns and other restrictions. In this paper, we consider the problem
of forecasting during such an event, and after it has occurred.

This is a different problem from change point detection. In the
situations we consider, we know that a change has occurred, and we want
to forecast the future after the change. Change point detection is about
identifying \emph{when} a change has occurred.

Time series models usually assume that the data evolve in the future in
a similar way to how they have evolved in the past. But a big event
(such as COVID-19) can result in a future that is different from the
past, at least in the short term.

The disruption may result in relatively simple changes in the series;
for example a level shift at the start of the disruption, and another at
the end of the disruption. Or they may be more complex, with changes to
the seasonal patterns, and changes to the level, which evolve over time.
In this paper, we consider a range of models that can be used to handle
such changes, and compare their performance on some real data sets.

We introduce several approaches to handling interruptions when
forecasting in Section~\ref{sec-methods}. We then apply these approaches
to two real data sets in \textbf{?@sec-sec-examples}. Finally, in
\textbf{?@sec-conclusions}, we discuss some of the advantages and
disadvantages of each approach, and provide some recommendations for
practitioners.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  11pt,
  a4paper,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[top=2.5cm,bottom=2.5cm,left=2.5cm,right=2.5cm]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{2}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage[section]{placeins}
\usepackage{orcidlink}
\definecolor{mypink}{RGB}{219, 48, 122}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[style=authoryear-comp,]{biblatex}
\addbibresource{references.bib}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Forecasting interrupted time series},
  pdfauthor={Rob J Hyndman; Bahman Rostami-Tabar},
  pdfkeywords={Forecasting, interrupted time series, disruptive events,
COVID-19},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

%% CAPTIONS
\usepackage{caption}
\DeclareCaptionStyle{italic}[justification=centering]
 {labelfont={bf},textfont={it},labelsep=colon}
\captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
\captionsetup[table]{style=italic,format=hang,singlelinecheck=true}

%% FONT
\usepackage{bera}
\usepackage[charter]{mathdesign}
\usepackage[scale=0.9]{sourcecodepro}
\usepackage[lf,t]{FiraSans}

%% HEADERS AND FOOTERS
\usepackage{fancyhdr}
\pagestyle{fancy}
\rfoot{\Large\sffamily\raisebox{-0.1cm}{\textbf{\thepage}}}
\makeatletter
\lhead{\textsf{\expandafter{\@title}}}
\makeatother
\rhead{}
\cfoot{}
\setlength{\headheight}{15pt}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}
\fancypagestyle{plain}{%
\fancyhf{} % clear all header and footer fields
\fancyfoot[C]{\sffamily\thepage} % except the center
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}}

%% MATHS
\usepackage{bm,amsmath}
\allowdisplaybreaks

%% GRAPHICS
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.8}

%% SECTION TITLES
\usepackage[compact,sf,bf]{titlesec}
\titleformat{\section}[block]
  {\fontsize{15}{17}\bfseries\sffamily}
  {\thesection}
  {0.4em}{}
\titleformat{\subsection}[block]
  {\fontsize{12}{14}\bfseries\sffamily}
  {\thesubsection}
  {0.4em}{}
\titlespacing{\section}{0pt}{*5}{*1}
\titlespacing{\subsection}{0pt}{*2}{*0.2}

%% BIBLIOGRAPHY.

\makeatletter
\@ifpackageloaded{biblatex}{
\ExecuteBibliographyOptions{bibencoding=utf8,minnames=1,maxnames=3, maxbibnames=99,dashed=false,terseinits=true,giveninits=true,uniquename=false,uniquelist=false,doi=false, isbn=false,url=true,sortcites=false}
\DeclareFieldFormat{url}{\texttt{\url{#1}}}
\DeclareFieldFormat[article]{pages}{#1}
\DeclareFieldFormat[inproceedings]{pages}{\lowercase{pp.}#1}
\DeclareFieldFormat[incollection]{pages}{\lowercase{pp.}#1}
\DeclareFieldFormat[article]{volume}{\mkbibbold{#1}}
\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
\DeclareFieldFormat[article]{title}{\MakeCapital{#1}}
\DeclareFieldFormat[article]{url}{}
\DeclareFieldFormat[inproceedings]{title}{#1}
\DeclareFieldFormat{shorthandwidth}{#1}
\usepackage{xpatch}
\xpatchbibmacro{volume+number+eid}{\setunit*{\adddot}}{}{}{}
% Remove In: for an article.
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{%
  \printtext{\bibstring{in}\intitlepunct}}}
\AtEveryBibitem{\clearfield{month}}
\AtEveryCitekey{\clearfield{month}}
\DeclareDelimFormat[cbx@textcite]{nameyeardelim}{\addspace}
\renewcommand*{\finalnamedelim}{\addspace\&\space}
}{}
\makeatother% Placement of logos

% Title and date

\title{Forecasting interrupted time series}
\date{19 February 2024}

\def\Date{\number\day}
\def\Month{\ifcase\month\or
 January\or February\or March\or April\or May\or June\or
 July\or August\or September\or October\or November\or December\fi}
\def\Year{\number\year}

% Working paper number and JEL codes

\makeatletter
\def\wp#1{\gdef\@wp{#1}}\def\@wp{??/??}
\def\jel#1{\gdef\@jel{#1}}\def\@jel{??}
\def\showjel{{\large\textsf{\textbf{JEL classification:}}~\@jel}}
\def\nojel{\def\showjel{}}
\makeatother

% Title page

\makeatletter
\def\cover{{\sffamily\setcounter{page}{0}
        \thispagestyle{empty}
        \placefig{2}{1.5}{width=5cm}{monash2}
        \placefig{16.9}{1.5}{width=2.1cm}{MBSportrait}
        \begin{textblock}{4}(16.9,4)ISSN 1440-771X\end{textblock}
        \begin{textblock}{7}(12.7,27.9)\hfill
        \includegraphics[height=0.7cm]{AACSB}~~~
        \includegraphics[height=0.7cm]{EQUIS}~~~
        \includegraphics[height=0.7cm]{AMBA}
        \end{textblock}
        \vspace*{2.5cm}
        \begin{center}\Large
        Department of Econometrics and Business Statistics\\[.5cm]
        \footnotesize http://monash.edu/business/ebs/research/publications
        \end{center}\vspace{2cm}
        \begin{center}
        \fbox{\parbox{14cm}{\begin{onehalfspace}\centering\Huge\vspace*{0.3cm}
                \textsf{\textbf{\expandafter{\@title}}}\vspace{1cm}\par
                \LARGE
                \expandafter{\@author}
                \end{onehalfspace}
        }}
        \end{center}
        \vfill
                \begin{center}\Large
                \Month~\Year\\[1cm]
                Working Paper \@wp
        \end{center}\vspace*{2cm}}}
        \def\addresses#1{\gdef\@addresses{#1}}\def\@addresses{??}
        \def\pageone{{\sffamily\setstretch{1}%
        \thispagestyle{empty}%
        \vbox to \textheight{%
        \raggedright\baselineskip=1.2cm
     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}
        \vspace{2cm}\par
        \hspace{1cm}\parbox{14cm}{\sffamily\large\@addresses}\vspace{1cm}\vfill
        \hspace{1cm}{\large\Date~\Month~\Year}\\[1cm]
        \hspace{1cm}\showjel\vss}}}
\def\blindtitle{{\sffamily
     \thispagestyle{plain}\raggedright\baselineskip=1.2cm
     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}\vspace{1cm}\par
        }}
\def\titlepage{{\cover\newpage\pageone\newpage\blindtitle}}

@@

% Authors

\nocover
  \author{Rob J Hyndman, Bahman Rostami-Tabar}
  \addresses{%
    %
      \textbf{Rob J Hyndman}\\%
      %
        Department of Econometrics \& Business Statistics\\%
        %
        %
        %
      %
        %
        Monash University\\%
        %
        %
      %
        %
        %
        Clayton VIC \\%
        %
      %
        %
        %
        %
        Australia\\%
      %
        %
        %
        %
        %
      %
      {Email: Rob.Hyndman@monash.edu}\\%
      \textit{Corresponding author}\\[0.5cm]%
   %
      \textbf{Bahman Rostami-Tabar}\\%
      %
        %
        Cardiff Business School\\%
        %
        %
      %
        %
        %
        %
        %
      %
        %
        %
        %
        United Kingdom\\%
      %
      {Email: rostami-tabarb@cardiff.ac.uk}\\%
      \\[0.5cm]%
   %
   }%
   \lfoot{\sf Hyndman, Rostami-Tabar: 19 February 2024}

% Keywords

\newenvironment{keywords}{\par\vspace{0.5cm}\noindent{\sffamily\textbf{Keywords:}}}{\vspace{0.25cm}\par\hrule\vspace{0.5cm}\par}

% Abstract
\renewenvironment{abstract}{\begin{minipage}{\textwidth}\parskip=1.4ex\noindent
\hrule\vspace{0.1cm}\par{\sffamily\textbf{\abstractname}}\newline\setstretch{1.5}}
  {\end{minipage}}
\begin{document}
\maketitle

\begin{abstract}
Forecasting interrupted time series data is a major challenge for
forecasting teams, especially in light of events such as the COVID-19
pandemic. This paper investigates several strategies for dealing with
interruptions in time series forecasting, including highly adaptable
models, intervention models, marking interrupted periods as missing,
forecasting what may have been, downweighting the interruption period,
and ensemble models. Each approach offers specific advantages and
disadvantages, such as adaptability, memory retention, data integrity,
flexibility, and accuracy. We evaluate the effectiveness of these
strategies using two actual datasets that were interrupted by COVID-19,
and we provide recommendations for how to handle these interruptions.
This work contributes to the literature on time series forecasting,
offering insights for academics and practitioners dealing with
interrupted data in numerous domains.
\end{abstract}

\begin{keywords}
  Forecasting, interrupted time series, disruptive events, COVID-19; 
  Forecasting, interrupted time series, disruptive events, COVID-19
\end{keywords}

\setstretch{1.5}
\section{Introduction}\label{introduction}

Time series forecasting models use historical data to estimate future
values \autocite{fildes2008forecasting} based on consistent patterns
observed in the past. However, time series are sometimes disrupted by
unusual events that jeopardise the regularity of the time series
pattern. This can be defined as an \emph{interrupted time series}. Time
series data may be disrupted by a variety of reasons, ranging from
temporary policy changes to natural disasters. Outbreaks, epidemics, and
pandemics, for example, may have a considerable impact on population
behavior, causing discrepancies and interruptions of data in a variety
of industries such as healthcare, pharmaceuticals, transportation,
retail, tourism, and traffic management. Similarly, supply chain
disruptions induced by factors such as strikes or shortages cause
irregularities in sales, inventory levels, production, and delivery,
often resulting in data gaps. Transportation-related data may be
interrupted due to equipment failures, such as malfunctions or
breakdowns in vehicles, aircraft, or rail systems, resulting in missing
or inaccurate data points. Furthermore, cybersecurity breaches data
quality by possibly skewing readings in datasets relating to financial
transactions, network traffic, and user activity, complicating time
series analysis.

The disruptions may result in relatively simple changes in the series;
for example a level shift at the start of the disruption, and another at
the end of the disruption. Or they may be more complex, with changes to
the seasonal patterns, and changes to the level, which evolve over time.

The presence of such disrupted events poses a significant challenge to
time series forecasting approaches, restricting their ability to
reliably capture systematic information and forecast future values. The
implications of these temporal interruptions are far-reaching: time
series forecasting techniques, which are a key component of the
forecaster's toolbox, may no longer be viable options since time series
models usually assume that the data will evolve in the future in a
similar way to how it has evolved in the past. But a big event (such as
COVID-19) can result in a future that is different from the past, at
least in the short term. Therefore, practitioners face a critical
question: how to effectively forecast time series data that are impacted
by such disruptive events? Many forecasters have faced this issue
recently with the COVID-19 pandemic, where historical patterns have been
severely disrupted due to policy interventions, lockdowns, and other
restrictions. This study is motivated by numerous conversations with
academics and practitioners, which demonstrate widespread concern about
dealing with data affected by events like the COVID-19 pandemic. In this
paper, we consider a range of models that can be used to forecast time
series influenced by disrupted events, and compare their performance on
some real data sets.

It is important to note that in this study, we consider the problem of
forecasting interrupted time series both during and after the event.
This is a separate problem from change point detection
\autocite{truong2020selective} and anomaly detection
\autocite{talagala2020anomaly}. In the situations we investigate, we
know that a change has occurred which has caused some unusual
observations, and we want to forecast what will happen next. Change
point detection refers to determining \emph{when} a change occurred
\autocite{blazquez2021review}, and anomaly detection aims to identify
unusual observations with the intention of either minimising their
value, or paying particular attention to them, and doing a root-cause
investigation. Our study focus also differs from \emph{intervention
modeling}, sometimes referred to as \emph{interrupted time series
analysis and modelling} \autocite{mcdowall2019interrupted}, which is
used when we have data about an outcome over time and the aim is to
assess the impact of an intervention, policy, or program implemented at
a specific time point. We are not trying to measure the impact of the
interruption; rather we are trying to produce sensible forecasts during
and after the interruption.

To our knowledge, this is the first study to describe and compare
general strategies for forecasting interrupted time series, such as
COVID-19 affected data. The study proposes and evaluates multiple
forecasting strategies designed to handle interruptions in time series
forecasting caused by disruptive events. These approaches include a
range of models capable of capturing different types of changes, from
simple level shifts to more complex alterations in seasonal patterns and
trends. By comparing the performance of these models on two real data
sets affected by the COVID-19 pandemic, the paper offers valuable
insights into the effectiveness and practical implications of each
approach. By discussing the advantages and disadvantages of each
strategy, and providing recommendations based on the findings, the paper
equips practitioners with actionable insights for addressing the
challenges associated with forecasting during and after disruptive
events. Further, by adhering to reproducibility principles and providing
both data and R code for the forecasting models, the research
contributes to transparency and promotes the generalizability of the
suggested strategies to a variety of domains. This openness allows for
the replication of results, ensuring the reliability of findings, and
enhancing the accessibility and utility of study outputs for the larger
research community.

We introduce several strategies to handling interruptions when
forecasting in Section~\ref{sec-methods}. We then apply these approaches
to two real data sets in \textbf{?@sec-sec-examples}. Finally, in
\textbf{?@sec-conclusions}, we discuss some of the advantages and
disadvantages of each approach, and provide some recommendations for
practitioners.

\section{Related works}\label{related-works}

The presence of unusual observations in time series analysis is a
significant concern that has received attention in numerous studies. In
this section, we provide a brief overview of some of these relevant
publications.

One modelling approach that has gained popularity is the intervention
model, also known as interrupted time series (ITS) modelling, introduced
by \textcite{box1975intervention}. This approach focuses on
understanding how and whether outcomes change following the
implementation of an intervention, policy, or programme at a specific
time point. ITS models provide a robust tool for analysing policy and
programme evaluation across diverse domains
\autocite{bernal2017interrupted,mcdowall2019interrupted}. These models
can forecast future values by developing models that incorporate a range
of components, such as pre-intervention levels, trends, seasonality,
covariates related to the intervention time, and other external factors
that might affect the outcome variable. A related approach is the
multiple regimes model, which is particularly useful in forecasting due
to its ability to capture structural changes and transitions observed in
interrupted time series data \autocite{Koopg2007}.

There are also statistical models that are used to detect changes in the
time series data and that can model unusual observation and have proven
effective in similar situations. A range of methods such as simple
location shift and scale shift models \autocite{aue2013structural},
Bayesian \autocite{barry1993bayesian}, and supervised and supervised
machine learning \autocite{aminikhanghahi2017survey}. Facebook's Prophet
\autocite{taylor2018forecasting} is an example of a time series
forecasting method that can incorporate changepoint detection
(automatically or manually), and adjust model parameters to capture the
newest trends in a time series. Anomaly detection is another related
area in time series analysis. In the context of anomaly detection, time
series anomalies are defined as points or subsequences with unexpected
or abnormal values, and the detection goal is to identify such anomalies
\autocite{blazquez2021review}.

In the context of forecasting disruptions such as the COVID-19 pandemic,
Bayesian approaches have been proposed for detecting changing points in
outbreaks and forecasting case numbers under counterfactual scenarios
\autocite{dehning2020inferring}. Counterfactual analysis examines what
would have happened to the forecast variable of interest if the
disruptive event had not occurred.
\textcite{athanasopoulos2023probabilistic} proposed a variant of this
approach for forecasting tourism recovery from COVID-19, combining
forecast reconciliation and forecast combinations applied to historical
data, to generate COVID-free counterfactual forecasts of what might have
been if the pandemic never occurred. Then scenario-based judgemental
probabilistic forecasts were compared with the counterfactual forecasts,
to better understand the future recovery of the tourism sector from the
pandemic.
